### 2008-10-08 trunk/Papers/rmrs-08/motivation.tex
changes to 2-3

\newcommand{\sem}[1]{\mathsf{#1}}
\newcommand{\sempred}[1]{\mathrm{#1}}

@@

We start with an informal discussion to 
illustrate how \rmrs\ can be used to represent partial
semantic information, and why people find it useful for this purpose.
Consider the following (toy) sentence:
\begin{examples}
  \item Every fat cat chased some dog.
\end{examples}

This sentence exhibits several kinds of ambiguity, including a scope
ambiguity between ``every fat cat'' and ``some dog'' and lexical
ambiguities of the nouns ``cat'' and ``dog'' (which have 8 and 7 
WordNet senses respectively).  Simplifying slightly by
ignoring tense information (but see
Section~\ref{sec:extensions}), two of these readings are
shown as logical forms below; they can be represented as trees as
shown in Fig.~\ref{fig:1}.

@@

Now imagine trying to extract semantic information from
the output of a part-of-speech tagger by using the word lemmas as
lexical predicate symbols.  Such a semantic representation is highly
partial, as it will not say anything about predicate-argument
relations or resolve lexical ambiguity---it will use predicate
symbols such as $\sempred{\_cat\_n}$, which might resolve to the
predicate symbols $\sem{\_cat\_n\_1}$ or $\sem{\_cat\_n\_2}$ in the
complete semantic representation.  (Notice the different
fonts for the ambiguous and unambiguous predicate symbols.)  
We can
express this information in \rmrs\ as follows.

@@

This \rmrs\ consists of six \emph{elementary predications} ({\sc ep}s), one
for each word lemma in the sentence; each of them is prefixed by a
\emph{label} and an \emph{anchor}, which are essentially variables
that refer to nodes in the trees in Fig.~\ref{fig:1}.  
The lack of semantic dependencies rendered by a {\sc pos} tagger
means we have unique arguments to each
predicate symbol (e.g., see $x_1$ and $x_3$ in (\ref{ex:cat-pos}),
compared with the {\em same} variable $x$ in the semantic
representations (\ref{ex:fat-cat-1}) and (\ref{ex:fat-cat-2})).
We can say that
the two trees \emph{satisfy} this \rmrs\ because it is possible to map
the labels and anchors into nodes in each tree and variable names like
$x_1$ and $x_3$ into variable names in the tree in such a way that the
predication
of
the node that labels and anchors are mapped to is consistent with
that given in the {\sc ep}s of the {\sc rmrs}.
For instance, the first tree satisfies
the \rmrs\ because $l_1$ and $a_1$ can map to the root of that
tree and $x_1$ to $x$, and the root label
$\sem{\_every\_q\_1}$ is consistent with the 
{\sc ep} predicate $\sempred{\_every\_q}$.

@@

This \rmrs\ uses two new types of atoms.  Atoms of the form $x_3=x_4$
express that the two variables $x_3$ and $x_4$ must map to the same
variable in any fully specific logical form; e.g., to the variable $x$
in Fig.~\ref{fig:1}.  Atoms of the form
$\ARG_i(a,z)$ and $\ARG_i(a,h)$ express that the $i$-th child
(counting from 0) of the
node to which the anchor $a$ refers is the variable name that $z$ denotes
(or the node that the {\em hole} $h$ denotes).  The separation of
{\sc ep}s and $\ARG$ atoms is perhaps the most important difference between
\rmrs\ and earlier underspecification formalisms such as \mrs\ and
dominance graphs, which required that a predicate and its arguments
had to be specified together.  \rmrs\ goes beyond these formalisms in
expressive power, because it allows one to state separately
information about how many arguments a predicate takes (which is
necessary for supporting parsers where information about lexical
subcategorisation is absent), and the values of
those (non-scopal and scopal) arguments.   If we also allow atoms of the form
$\ARG_{\{2,3\}}(a,x)$ to express uncertainty as to whether $x$ is the
second or third child of the anchor $a$, then \rmrs\ can specify the
arguments to a predicate while underspecifying their position.

Finally, \rmrs\ (\ref{ex:cat-erg}) is a notational variant of the {\sc
  mrs} derived by the {\sc erg}, a wide-coverage deep grammar:
\begin{examples}
\item $l_1:a_1\handel\sempred{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mathsf{RSTR}(a_1,h_2),
\mathsf{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\sempred{\_fat\_j\_1}(e'), \ARG_1(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\sempred{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\sempred{\_chase\_v\_1}(e)$,\\
\hspace*{0.1in}$\ARG_1(a_5,x_4),
\ARG_2(a_5,x_5)$\\ 
$l_6:a_6\handel\sempred{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mathsf{RSTR}(a_6,h_7),
\mathsf{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\sempred{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\label{ex:cat-erg}
\end{examples}
$\mathsf{RSTR}$ and $\mathsf{BODY}$ are conventional names for the
scopal arguments to predicate symbols corresponding to quantifiers.
Atoms like $h_2 \qeq l_{42}$ specify a kind of ``outscopes''
relationship between the hole and the label, and can be used to
underspecify the scope of the two quantifiers.  Notice also that the
labels of the {\sc ep}s for ``fat'' and ``cat'' are stipulated to be
equal in the \rmrs\ (\ref{ex:cat-erg}), whereas the anchors are not.
In the tree, it is the anchors that are mapped to the nodes with the
labels $\sem{\_fat\_j\_1}$ and $\sem{\_cat\_n\_1}$; the label is
mapped to the conjunction node just above them.  In other words, the
role of the anchor in an {\sc ep} is to connect a predicate to its
arguments, while the role of the label is to connect the {\sc ep} to
the surrounding formula.  Representing conjunction with label sharing
stems from \mrs\ and provides compact semantic representations.
Finally, (\ref{ex:cat-erg}) intuitively represents more specific
semantic information than (\ref{ex:cat-pos}), so long as the
constructors that \_cat\_n\_1 can correspond to is a subset of those
that \_cat\_n can correspond to (written
$\sempred{\_cat\_n\_1}\sqsubseteq \sempred{\_cat\_n}$), and similarly
for \_fat\_j\_1 etc.  This is because with these so-called SPEC
relations on the predicate symbols in place, (\ref{ex:cat-erg}) simply
adds ARGs, qeqs and variable equalities to (\ref{ex:cat-pos}).

In summary, \rmrs\ is a formalism for representing partial information
about semantic representations.  It can
represent the output of a deep grammar in a way that is a notational
variant of \mrs; but in cases in which a deep grammar is not
available, it can be used to capture whatever partial semantic
information a shallower {\sc nlp} tool can provide.  \rmrs\ is
designed to be a generic formalism for partial semantic
representations which abstracts over the details of different {\sc nlp}
tools and allows us to represent, process, and compare information
from different tools in the same formalism, and in the long run
provide a platform for incrementally enriching partial information
such as (\ref{ex:cat-pos}) in order to eventually approximate more
detailed representations such as (\ref{ex:cat-partial-parser}) or
(\ref{ex:cat-erg}).

However, although people have started using \rmrs\ in their
implemented systems, it is still relatively ad-hoc in that its formal
syntax is not consistent across different papers, and its formal
semantics is not defined; we don't know, formally, what an \rmrs\
\emph{means} in terms of semantic representations like
(\ref{ex:fat-cat-1}) and (\ref{ex:fat-cat-2}), and this hinders our
ability to design efficient algorithms for processing \rmrs. The
purpose of this paper is to lay the groundwork for fixing this problem.

>>>>

@@

We start with an informal illustration of why \rmrs\ is necessary and
useful to represent partial semantic information.  Consider the
following (toy) sentence:

\begin{examples}
  \item Every fat cat chased some dog.
\end{examples}

This sentence exhibits several kinds of ambiguity, including a scope
ambiguity between ``every fat cat'' and ``some dog'' and lexical
ambiguities of the nouns ``cat'' and ``dog'' (which have 8 and 7
WordNet senses respectively).  Simplifying slightly by ignoring tense
information, two of these readings are shown as logical forms below;
these can be represented as trees as shown in Fig.~\ref{fig:1}.
%  (but see Section~\ref{sec:extensions})

@@

Now imagine we are trying to extract semantic information from the
output of a part-of-speech tagger by using the word lemmas as lexical
predicate symbols.  Such a semantic representation is highly partial,
as it will not say anything about predicate-argument relations or
resolve lexical ambiguity---it will use predicate symbols such as
$\sempred{\_cat\_n}$, which might resolve to the predicate symbols
$\sem{\_cat\_n\_1}$ or $\sem{\_cat\_n\_2}$ in the complete semantic
representation.  (Notice the different fonts for the ambiguous and
unambiguous predicate symbols.)

Partial semantic information, especially with respect to ambiguity, is
typically represented using an underspecification formalism such as
\mrs\ \cite{copestake:etal:2005} or CLLS \cite{egg:etal:inpress}.  But
most underspecification formalisms are unable to represent semantic
information that is as partial as what we get from a POS tagger: They
assume that the complete predicate-argument structure is known and
there is no lexical ambiguity.  \rmrs\ \cite{copestake:2007a} was
designed to address this problem.  In \rmrs, we can express the
information we get from the POS tagger as follows:

@@

This \rmrs\ expresses only that certain predications must be made in
the semantic representation -- but it doesn't say anything about the
structure of the semantic representation, about most arguments of the
predicates (note that $\sempred{\_chase\_v}(e)$ doesn't say who chases
whom), or about the coindexation of variables ($\sempred{\_every\_q}$
binds the variable $x_1$, whereas $\sempred{\_cat\_n}$ speaks about
$x_3$), and maintains the lexical ambiguities.  Technically, it
consists of six \emph{elementary predications} ({\sc ep}s), one for
each word lemma in the sentence; each of them is prefixed by a
\emph{label} and an \emph{anchor}, which are essentially variables
that refer to nodes in the trees in Fig.~\ref{fig:1}.  We can say that
the two trees \emph{satisfy} this \rmrs\ because it is possible to map
the labels and anchors into nodes in each tree and variable names like
$x_1$ and $x_3$ into variable names in the tree in such a way that the
predication of the node that labels and anchors are mapped to is
consistent with that given in the {\sc ep}s of the {\sc rmrs}.  For
instance, the first tree satisfies the \rmrs\ because $l_1$ and $a_1$
can map to the root of that tree and $x_1$ to $x$, and the root label
$\sem{\_every\_q\_1}$ is consistent with the {\sc ep} predicate
$\sempred{\_every\_q}$.

% The lack of semantic dependencies rendered by a {\sc pos} tagger means
% we have unique arguments to each predicate symbol (e.g., see $x_1$ and
% $x_3$ in (\ref{ex:cat-pos}), compared with the {\em same} variable $x$
% in the semantic representations (\ref{ex:fat-cat-1}) and
% (\ref{ex:fat-cat-2})).

@@

This \rmrs\ uses two new types of atoms.  Atoms of the form $x_3=x_4$
express that the two variables $x_3$ and $x_4$ must map to the same
variable in any fully specific logical form; e.g., both to the
variable $x$ in Fig.~\ref{fig:1}.  Atoms of the form $\ARG_i(a,z)$ and
$\ARG_i(a,h)$ express that the $i$-th child (counting from 0) of the
node to which the anchor $a$ refers is the variable name that $z$
denotes (or the node that the {\em hole} $h$ denotes).  So unlike
earlier underspecification formalisms, \rmrs\ can specify the
predicate of an atom separately from its (scopal and non-scopal)
arguments; this is necessary for supporting parsers where information
about lexical subcategorisation is absent). If we also allow atoms of
the form $\ARG_{\{2,3\}}(a,x)$ to express uncertainty as to whether
$x$ is the second or third child of the anchor $a$, then \rmrs\ can
even specify the arguments to a predicate while underspecifying their
position.  This is useful for \todo{what?}.

Finally, the \rmrs\ (\ref{ex:cat-erg}) is a notational variant of the
{\sc mrs} derived by the {\sc erg}, a wide-coverage deep grammar:
\begin{examples}
\item $l_1:a_1\handel\sempred{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mathsf{RSTR}(a_1,h_2),
\mathsf{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\sempred{\_fat\_j\_1}(e'), \ARG_1(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\sempred{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\sempred{\_chase\_v\_1}(e)$,\\
\hspace*{0.1in}$\ARG_1(a_5,x_4),
\ARG_2(a_5,x_5)$\\ 
$l_6:a_6\handel\sempred{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mathsf{RSTR}(a_6,h_7),
\mathsf{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\sempred{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\label{ex:cat-erg}
\end{examples}

Here we have full information about the predicate-argument structure
and the variable binding; we only underspecify scope and some lexical
ambiguity.  $\mathsf{RSTR}$ and $\mathsf{BODY}$ are conventional names
for the $\ARG_1$ and $\ARG_2$ of a quantifier predicate symbol.  Atoms
like $h_2 \qeq l_{42}$ (``qeq'') specify a certain kind of
``outscopes'' relationship between the hole and the label, and are
used here to underspecify the scope of the two quantifiers.

Notice that the labels of the {\sc ep}s for ``fat'' and ``cat'' are
stipulated to be equal in this \rmrs, whereas the anchors are not.  In
the tree, it is the anchors that are mapped to the nodes with the
labels $\sem{\_fat\_j\_1}$ and $\sem{\_cat\_n\_1}$; the label is
mapped to the conjunction node just above them.  In other words, the
role of the anchor in an {\sc ep} is to connect a predicate to its
arguments, while the role of the label is to connect the {\sc ep} to
the surrounding formula.  Representing conjunction with label sharing
stems from \mrs\ and provides compact semantic representations.

Finally, (\ref{ex:cat-erg}) uses predicate symbols like
$\sempred{\_dog\_n\_1}$ that are meant to be more specific than the
symbols like $\sempred{\_dog\_n}$ which the earlier \rmrs s used.
This reflects the fact that the deep grammar will perform some lexical
disambiguation, whereas the chunker and POS tagger didn't.  The fact
that the former symbol should be more specific than the latter can be
represented using SPEC atoms like $\sempred{\_dog\_n\_1}\sqsubseteq
\sempred{\_dog\_n}$), which we have omitted in (\ref{ex:cat-erg}) for
brevity.  Note that even a deep grammar will not fully disambiguate to
semantic predicate symbols, such as WordNet senses, and so
$\sempred{\_dog\_n\_1}$ can still be consistent with multiple symbols
like $\sem{\_dog\_n\_1}$ and $\sem{\_dog\_n\_2}$ in the semantic
representation.

In summary, \rmrs\ is a formalism for partial semantic
representations.  It allows us to represent the semantic information
that can be extracted from a wide range of NLP tools in a uniform way.
This is useful for hybrid systems which fall back to shallower
analyses if deeper parsing is unsuccessful, or which try to match
deeply parsed queries against shallow parses of large text corpora;
and in fact, \rmrs\ is gaining popularity as a practical interchange
format for exactly these purposes \cite{copestake:2003,frank:2004}.
However, \rmrs\ is still relatively ad-hoc in that its formal syntax
is not consistent across different papers, and its formal semantics is
not defined; we don't know, formally, what an \rmrs\ \emph{means} in
terms of semantic representations like (\ref{ex:fat-cat-1}) and
(\ref{ex:fat-cat-2}), and this hinders our ability to design efficient
algorithms for processing \rmrs. The purpose of this paper is to lay
the groundwork for fixing this problem.
